PREPRINT -- Discriminating Seagrasses From Green Macroalgae in European Intertidal areas using high resolution multispectral drone imagery -- PREPRINT.

### 2024-03-06 docs/_tex/index.tex
mistakes correction

Coastal areas are vital hotshots for marine biodiversity, with
intertidal seagrass meadows playing a crucial role at the interface
between land and ocean \citep{unsworth2022}. Seagrass meadows provide a
myriad of ecosystem services to humanity, including carbon
sequestration, oxygen production,protection against sea-level rise and
coastline erosion, and limitation of eutrophication. They serve as vital
habitats for a diverse array of marine and terrestrial species,
providing living, breeding, and feeding grounds \citetext{\citealp[
]{gardner2018}; \citealp[ ]{Zoffoli2022}; \citealp{jankowska2019}}. Due
to the concentration of human activities in coastal zones, seagrass
meadows are directly exposed to and impacted by anthropogenic pressures.
Global regression and fragmentation are currently observed due to
climate change, diseases, urbanization, land reclamation, dredging,
competition with alien species, and reduction in water quality
\citetext{\citealp[ ]{nguyen2021}; \citealp[ ]{soissons2018}; \citealp[
]{orth2006}; \citealp[ ]{lin2018}; \citealp{duffy2019}}. While
improvements in water quality have been recently reported in Europe,
allowing an overall recovery of seagrass ecosystems at local scale, many
coastal waters worldwide are still subjected to strong eutrophication
processes \citetext{\citealp[ ]{deSantos2019}; \citealp{Zoffoli2021}}.
Coastal eutrophication has been associated to anomalous accumulation of
green macroalgae, the so-called green tides (\citep{devlin2023nutrients}
; \citep{devlin2023nutrients}). Green tides produce shade and
suffication over seagrass individuals, thus threatening the health of
seagrass ecosystems \citep{wang2022}.

The importance of seagrass meadows and the variety of ecosystem services
they provide have led to the enhancement of both global and regional
programs monitoring Essential Oceanic Variable (EOVs) such as seagrass
composition \citep{Miloslavich2018}, as well as Essential Biodiversity
Variable (EBVs) such as seagrass taxonomic diversity, species
distribution, populuation abundance, and phenology \citep{Pereira2013}.
Seagrass monitoring is all the more important as it makes possible to
identify which pressure should be addressed in priority to facilitate
effective mitigation actions. Traditionally, indicators of seagrass
status have been quantified using in situ measurements. The acquisition
of field data in intertidal zones is however notoriously challenging.
Intertidal seagrass meadows are only partially exposed during low tide
and can be situated in difficult-to-reach mudflats, potentially leading
to inaccurate and limited estimations with conventional sampling
techniques \citep{nijland2019}. Satellite observations have been proven
effective in complementing in situ sampling, allowing for the near
real-time and consistent retrieval of seagrass EOVs and EBVs over
extensive meadows. \citetext{\citealp[ ]{Zoffoli2021}; \citealp[
]{xu2021}; \citealp[ ]{Traganos2018}; \citealp{coffer2023}}

While satellite remote sensing provides temporally consistent
observations over large spatial scales, it is limited by several
constraints. Satellite missions with a high temporal resolution
(e.g.~daily MODIS observation) are limited by a coarse spatial
resolution (\textgreater100m). Mission with a high spatial resolution
such as Sentinel-2 (10m) or Landsat8/9 (30m) can be limited by low
spectral resolution. The limited number of spectral bands is indeed a
challenge to accurately discriminate seagrass from others co-existing
macrophytes. In particular Chlorophyceae (green algae) and marine
Magnoliopsida (seagrass) share the same pigment composition
\citetext{\citealp[ ]{ralph2002}; \citealp{Douay2022}}. As a result,
their respective spectral signatures can be considered similar by a
non-expert observer \citetext{\citealp[
]{Davies2023}; \citealp{bannari2022}}. Recently, using advanced
machine-learning algorithms trained with a large hyperspectral library
of more then 300 field reflectance measurements, \citep{Davies2023}
demonstrated that it was possible to discriminate Magnoliopsida from
Chlorophyceae using radiometric data acquired at Sentinel-2 's spectral
resolution. The application of this approach to satellite remote sensing
remains however to be validated. Moreover patches of green algae can
develop at small spatial scales that are not observable using
non-commercial satellite imagery \citep{tuya2013}, especially during the
initial stage of the event.

Drone can potentially fill the gaps left by satellite remote sensing and
in situ measurments, due to their ability to provide spatially-explicit
observations at very high spatial resolutions (pixel size from cm to mm)
and still capturing data at multi-spectral resolution
\citetext{\citealp[ ]{fairley2022drone}; \citealp{oh2017use}}. The
versatility of drones allows for their application across a diverse
thematic range , from coastal zone management \citetext{\citealp[
]{adade2021}; \citealp[ ]{casella2020}; \citealp{angnuureng2022}} to
mapping species distribution \citetext{\citealp[ ]{joyce2023}; \citealp[
]{tallam2023}; \citealp[ ]{Roca2022}; \citealp[
]{Roman2021}; \citealp{Brunier2022Topographic}}. However, when applied
to coastal habitat mapping, most sutdy cases were limited to a single
flight, restricting the generalizability of their application over wider
geographical scales \citetext{\citealp[ ]{Roman2021}; \citealp[
]{collin2019improving}; \citealp[
]{rossiter2020uav}; \citealp{Brunier2022Topographic}}. The present study
aimed at analyzing the potential of multispectral drone remote sensing
to map intertidal macrophytes, with a particular focus on discriminating
Magnoliopsida and Chlorophyceae. Ten drone flights were performed over
soft-bottom intertidal areas along two European countries (France and
Portugal), covering a wide range of habitats, from monospecific seagrass
meadows to meadows mixed with green or red algae. A deep learning
algorithm was trained and validated for macrophyte discrimination,
emphasizing applicability across diverse sites without a loss of
prediction accuracy.

@@

Seven study sites distributed between France and Portugal were selected
for their relatively extensive intertidal seagrass beds. Two sites are
located in Gulf of Morbihan (Figure~\ref{fig-map} A), France (47.5791°N,
2.8018°W). This gulf covers an area of 115 km² and is only connect to
the sea through a 900m wide channel. A total of 53 small islands are
scattered across the gulf leading to 250 km of shorelines. Patchy
seagrass meadows can be found on a lot of these islands. Within the Gulf
of Morbihan we have chosen two sites : one is located in one of these
island (Arz island); the other one is located in the sourthern part of
the gulf (Duer). Two others sites are located in Bourgneuf Bay, France
(46.9849°N, 2.1488°W). This bay is a semi-enclosed macrotidal bay,
protected from waves by Noirmoutier Island. Bougneuf bay hosts a large
intertidal seagrass meadow of about 6 km². Within this meadow, the sites
observed by drones (L'Epine and Barbatre, Figure~\ref{fig-map} B)
contain a monospecific bed of \emph{Zostera noltei} with very little
mixing with other macrophytes. Three sites have been surveyed in the Ria
de Aveiro Lagoon in Portugal (40.6887°N, 8.6810°W). The extent of this
lagoon is about 80 km² with many narrow channels, large salt marshes and
many mudflats that uncovers at low tide. It is connected to the open sea
through a single channel, with a tidal lag between the North and the
South of the lagoon. The southernmost site (Gafanha) is a mudflat
located in an affluent of the lagoon whereas the two other sites are
situated in the middle of the lagoon and are only accessible by boat
(Figure~\ref{fig-map} C). These Portuguese sites are characterized by a
more diverse intertidal vegetation, where patches of seagrass could
intermingle with red algae (Rhodophyceae), brown algae (Phaeophyceae),
and green algae (Chlorophyceae).

@@

At each location, a DJI Matrice 200 quadcopter drone equipped with a
Micasense RedEdge Dual MX multispectral camera was flown to take 1.2
million pixel reflectance photographs in 10 spectral bands, from 444 nm
(blue) to 840 nm (near infrared, NIR). An angle of 90° was maintained
between the sun and the drone's route to guarantee uniform lighting
conditions between flight lines. A side overlap of 70\% and a front
overlap of 80\% between each imagewas set for each flight. A downwelling
light sensor (DLS2) was used to acquire irradiance data concomitantly
with the camera measurements. Raw data were calibrated in reflectance
using a calibration panel reflective at \textasciitilde50\% provided by
the manufacturer. A structure-from-motion photogrammetry software
(Agisoft Metashape) was used to process images to obtain multispectral
orthomosaics of each flight. The workflow for orthomosaicking was the
same for every flight. First, tying key points were detected inside of
each image and between overlapping images in order to obtain a sparse
point cloud. This cloud was cleaned using reprojection accuracy metric
in order to remove noisy points. A dense point cloud was then produced
using a structure from motion algorithm. A surface interpolation of this
dense point cloud was made to obtain a digital surface model (DSM), used
to reconstruct the multispectral ortho-image. Across all sites, flights
were made at two different altitudes : 12 m or/and 120 m. Low altitude
drone flights produce ortho-images with a very high spatial resolution
(8 mm per pixel), making it simple to visually distinguish between the
various types of vegetation. High altitude fights on the other hand
allow to cover large areas and produced images with a pixel size of 80
mm (Table 1).

@@

Before each flight, targets used as Ground Control Points (GCPs) were
distributed over the study site and georeferenced with a Trimble © Geo
XH 6000 differential GPS (dGPS). GCPs were used to correct
georeferencing imprecision of orthomosaics with an horizontal and
vertical accuracy of 10cm. dGPS was also used to georeference quadrats
of 0.25cm² used to assess the presence or absence of 5 key classes of
vegetation : Bacillariophyceae (Benthic diatoms forming large biofilm
over the sediment during low tide), Phaeophyceae (brown algae),
Magnoliopsida (seagrass), Chlorophyceae (green algae) and Rhodophyceae
(red algae) (Figure~\ref{fig-vegetation}). Pictures of each quadrat were
uploaded on the online on the Global Biodiversity Information Facility
(GBIF) platform \citep{BedeGbif}, a public open portal to store and
share biodiversity data. Each photograph was also processed to estimate
the percent cover of each type of vegetation using an image processing
software (imageJ). For all quadrat, the hyperspectral reflectance
signatures of each vegetation class was measured using an ASD FieldSpec
HandHeld 2 spectroradiometer, which acquires reflectance between 325 and
1075 nm, with 1 nm of spectral resolution. The methodological scheme
followed in this study is presented in Figure~\ref{fig-workflow}.

@@

\caption{\label{fig-VIP}Variable Importance of the Neural Network
Classifier for each vegetation class. The bigger the slice, the more
important is the variable to predict accuratly this class. The bottom
right plot shows the drone standardised reflectance spectra of each
class.}

@@

The primary objective of this study was to develop a method for the
accurate classification of macrophytes on intertidal mudflats,
specifically focusing on distinguishing between Chlorophyceae (green
algae) and marine Magnoliopsida (seagrasses) using multispectral drone
data. The discrimination of seagrasses from green macroalgae presents
significant challenges, primarily due to the shared pigment composition
between these taxa, as depicted in Figure~\ref{fig-Pigm} but also by the
frequent spatial mixing of these green macrophytes. This similarity in
pigment composition, including photosynthetic and carotenoid pigments
like chlorophyll b and c, fucoxanthin, and neoxanthin, results in
spectral signatures that are often indistinguishable to non exper eyes.
This spectral overlap is further complicated by the intermingling of
seagrasses and green macroalgae in the same spatial locations, posing a
significant challenge for remote sensing aimed at accurately mapping and
monitoring coastal ecosystems. Our study addresses these complexities by
utilizing high-resolution drone imagery with 10 spectral bands.
\citep{Davies2023} has shown that having at least 8 spectral bands
ranging between 500 nm to 850 nm with including a band at 530 nm and
another one at 730 nm is crucial to accurately discriminate green
macrophytes.

@@

Climate change, global warming, alien and invasive species development,
coastal erosion, sealevel rise, while further impact coastal ecosystems
in the future \citetext{\citealp{SCHIBALSKI2022101414}; \citealp[
]{holon2018predictive}; \citealp{marquet2024global}}. The demand for
meaningful and efficient monitoring methods for these habitats has
therefore never been greater \citetext{\citealp[
]{muller2018satellite}; \citealp[
]{villalobos2023remote}; \citealp{oiry2021using}}. Employing traditional
sampling methods to monitor these ecosystems is time and
resource-intensive, and the findings are often difficult to scale up.
Earth observation methods can bridge this gap and meet the needs for
monitoring coastal ecosystems. \citep{papathanasopoulou2019satellite}.
The retrieval of Essential Biodiversity Variables (EBVs) and Essential
Ocean Variables (EOVs) through satellite observations is increasingly
common, enabling comprehensive monitoring of entire ecosystems over
extended time periods \citetext{\citealp[
]{ratnarajah2023monitoring}; \citealp{ZOFFOLI2020112020}}. The
significance of monitoring waters, particularly in light of the Water
Framework Directive (WFD ; \citep{WFD2000}), cannot be overstated in
environmental management and conservation efforts. The WFD mandates the
achievement and maintenance of ``good ecological status'' for all
European waters, which necessitates a comprehensive understanding and
monitoring of aquatic ecosystems, including coastal habitats like
seagrass \citetext{\citealp[ ]{foden2007angiosperms}; \citealp[
]{nordlund2024one}; \citealp{Zoffoli2021}}.

>>>>

Coastal areas are vital hotspots for marine biodiversity, with
intertidal seagrass meadows playing a crucial role at the interface
between land and ocean \citep{unsworth2022}. Seagrass meadows provide a
myriad of ecosystem services to humanity, including carbon
sequestration, oxygen production, protection against sea-level rise and
coastline erosion, and limitation of eutrophication. They serve as vital
habitats for a diverse array of marine and terrestrial species,
providing living, breeding, and feeding grounds \citetext{\citealp[
]{gardner2018}; \citealp[ ]{Zoffoli2022}; \citealp{jankowska2019}}. Due
to the concentration of human activities in coastal zones, seagrass
meadows are directly exposed to and impacted by anthropogenic pressures.
Global regression and fragmentation are currently observed due to
climate change, diseases, urbanization, land reclamation, dredging,
competition with alien species, and reduction in water quality
\citetext{\citealp[ ]{nguyen2021}; \citealp[ ]{soissons2018}; \citealp[
]{orth2006}; \citealp[ ]{lin2018}; \citealp{duffy2019}}. While
improvements in water quality have been recently reported in Europe,
allowing an overall recovery of seagrass ecosystems at local scale, many
coastal waters worldwide are still subjected to strong eutrophication
processes \citetext{\citealp[ ]{deSantos2019}; \citealp{Zoffoli2021}}.
Coastal eutrophication has been ass ociated to anomalous accumulation of
green macroalgae, the so-called green tides
(\citep{devlin2023nutrients}). Green tides produce shade and suffocation
over seagrass individuals, thus threatening the health of seagrass
ecosystems \citep{wang2022}.

The importance of seagrass meadows and the variety of ecosystem services
they provide have led to the enhancement of both global and regional
programs monitoring Essential Oceanic Variable (EOVs) such as seagrass
composition \citep{Miloslavich2018}, as well as Essential Biodiversity
Variable (EBVs) such as seagrass taxonomic diversity, species
distribution, population abundance, and phenology \citep{Pereira2013}.
Seagrass monitoring is all the more important as it makes possible to
identify which pressure should be addressed in priority to facilitate
effective mitigation actions. Traditionally, indicators of seagrass
status have been quantified using in situ measurments. The acquisition
of field data in intertidal zones is however notoriously challenging.
Intertidal seagrass meadows are only partially exposed during low tide
and can be situated in difficult-to-reach mudflats, potentially leading
to inaccurate and limited estimations with conventional sampling
techniques \citep{nijland2019}. Satellite observations have been proven
effective in complementing in situ sampling, allowing for the near
real-time and consistent retrieval of seagrass EOVs and EBVs over
extensive meadows. \citetext{\citealp[ ]{Zoffoli2021}; \citealp[
]{xu2021}; \citealp[ ]{Traganos2018}; \citealp{coffer2023}}

While satellite remote sensing provides temporally consistent
observations over large spatial scales, it is limited by several
constraints. Satellite missions with a high temporal resolution
(e.g.~daily MODIS observation) are limited by a coarse spatial
resolution (\textgreater100m). Mission with a high spatial resolution
such as Sentinel-2 (10m) or Landsat8/9 (30m) can be limited by low
spectral resolution. The limited number of spectral bands is indeed a
challenge to accurately discriminate seagrass from others co-existing
macrophytes. In particular Chlorophyceae (green algae) and marine
Magnoliopsida (seagrass) share the same pigment composition
\citetext{\citealp[ ]{ralph2002}; \citealp{Douay2022}}. As a result,
their respective spectral signatures can be considered similar by a
non-expert observer \citetext{\citealp[
]{Davies2023}; \citealp{bannari2022}}. Recently, using advanced
machine-learning algorithms trained with a large hyperspectral library
of more then 300 field reflectance measurments, \citep{Davies2023}
demonstrated that it was possible to discriminate Magnoliopsida from
Chlorophyceae using radiometric data acquired at Sentinel-2 's spectral
resolution. The application of this approach to satellite remote sensing
remains however to be validated. Moreover patches of green algae can
develop at small spatial scales that are not observable using
non-commercial satellite imagery \citep{tuya2013}, especially during the
initial stage of the event.

Drones can potentially fill the gaps left by satellite remote sensing
and in situ measurments, due to their ability to provide
spatially-explicit observations at very high spatial resolutions (pixel
size from cm to mm) and still capturing data at multi-spectral
resolution \citetext{\citealp[
]{fairley2022drone}; \citealp{oh2017use}}. The versatility of drones
allows for their application across a diverse thematic range , from
coastal zone management \citetext{\citealp[ ]{adade2021}; \citealp[
]{casella2020}; \citealp{angnuureng2022}} to mapping species
distribution \citetext{\citealp[ ]{joyce2023}; \citealp[
]{tallam2023}; \citealp[ ]{Roca2022}; \citealp[
]{Roman2021}; \citealp{Brunier2022Topographic}}. However, when applied
to coastal habitat mapping, most study cases were limited to a single
flight, restricting the generalizability of their application over wider
geographical scales \citetext{\citealp[ ]{Roman2021}; \citealp[
]{collin2019improving}; \citealp[
]{rossiter2020uav}; \citealp{Brunier2022Topographic}}. The present study
aimed at analyzing the potential of multispectral drone remote sensing
to map intertidal macrophytes, with a particular focus on discriminating
Magnoliopsida and Chlorophyceae. Ten drone flights were performed over
soft-bottom intertidal areas along two European countries (France and
Portugal), covering a wide range of habitats, from monospecific seagrass
meadows to meadows mixed with green or red algae. A deep learning
algorithm was trained and validated for macrophyte discrimination,
emphasizing applicability across diverse sites without a loss of
prediction accuracy.

@@

Seven study sites distributed between France and Portugal were selected
for their relatively extensive intertidal seagrass beds. Two sites are
located in Gulf of Morbihan (Figure~\ref{fig-map} A), France (47.5791°N,
2.8018°W). This gulf covers an area of 115 km² and is only connected to
the sea through a 900m wide channel. A total of 53 small islands are
scattered across the gulf leading to 250 km of shorelines. Patchy
seagrass meadows can be found on a lot of these islands. Within the Gulf
of Morbihan we have chosen two sites : one is located in one of these
island (Arz island); the other one is located in the southern part of
the gulf (Duer). Two others sites are located in Bourgneuf Bay, France
(46.9849°N, 2.1488°W). This bay is a semi-enclosed macrotidal bay,
protected from waves by Noirmoutier Island. Bougneuf bay hosts a large
intertidal seagrass meadow of about 6 km². Within this meadow, the sites
observed by drones (L'Epine and Barbatre, Figure~\ref{fig-map} B)
contain a monospecific bed of \emph{Zostera noltei} with very little
mixing with other macrophytes. Three sites have been surveyed in the Ria
de Aveiro Lagoon in Portugal (40.6887°N, 8.6810°W). The extent of this
lagoon is about 80 km² with many narrow channels, large salt marshes and
many mudflats that uncovers at low tide. It is connected to the open sea
through a single channel, with a tidal lag between the North and the
South of the lagoon. The southernmost site (Gafanha) is a mudflat
located in an affluent of the lagoon whereas the two other sites are
situated in the middle of the lagoon and are only accessible by boat
(Figure~\ref{fig-map} C). These Portuguese sites are characterized by a
more diverse intertidal vegetation, where patches of seagrass could
intermingle with red algae (Rhodophyceae), brown algae (Phaeophyceae),
and green algae (Chlorophyceae).

@@

At each location, a DJI Matrice 200 quadcopter drone equipped with a
Micasense RedEdge Dual MX multispectral camera was flown to take 1.2
million pixel reflectance photographs in 10 spectral bands, from 444 nm
(blue) to 840 nm (near infrared, NIR). An angle of 90° was maintained
between the sun and the drone's route to guarantee uniform lighting
conditions between flight lines. A side overlap of 70\% and a front
overlap of 80\% between each image was set for each flight. A
downwelling light sensor (DLS2) was used to acquire irradiance data
concomitantly with the camera measurements. Raw data were calibrated in
reflectance using a calibration panel reflective at \textasciitilde50\%
provided by the manufacturer. A structure-from-motion photogrammetry
software (Agisoft Metashape) was used to process images to obtain
multispectral orthomosaics of each flight. The workflow for
orthomosaicking was the same for every flight. First, tying key points
were detected inside of each image and between overlapping images in
order to obtain a sparse point cloud. This cloud was cleaned using
reprojection accuracy metric in order to remove noisy points. A dense
point cloud was then produced using a structure from motion algorithm. A
surface interpolation of this dense point cloud was made to obtain a
digital surface model (DSM), used to reconstruct the multispectral
ortho-image. Across all sites, flights were made at two different
altitudes : 12 m or/and 120 m. Low altitude drone flights produce
ortho-images with a very high spatial resolution (8 mm per pixel),
making it simple to visually distinguish between the various types of
vegetation. High altitude fights on the other hand allow to cover large
areas and produced images with a pixel size of 80 mm (Table 1).

@@

Before each flight, targets used as Ground Control Points (GCPs) were
distributed over the study site and georeferenced with a Trimble © Geo
XH 6000 differential GPS (dGPS). GCPs were used to correct
georeferencing imprecision of orthomosaics with an horizontal and
vertical accuracy of 10cm. dGPS was also used to georeference quadrats
of 0.25 m² used to assess the presence or absence of 5 key classes of
vegetation : Bacillariophyceae (Benthic diatoms forming large biofilm
over the sediment during low tide), Phaeophyceae (brown algae),
Magnoliopsida (seagrass), Chlorophyceae (green algae) and Rhodophyceae
(red algae) (Figure~\ref{fig-vegetation}). Pictures of each quadrat were
uploaded online on the Global Biodiversity Information Facility (GBIF)
platform \citep{BedeGbif}, a public open portal to store and share
biodiversity data. Each photograph was also processed to estimate the
percent cover of each type of vegetation using an image processing
software (imageJ). For all quadrat, the hyperspectral reflectance
signatures of each vegetation class was measured using an ASD FieldSpec
HandHeld 2 spectroradiometer, which acquires reflectance between 325 and
1075 nm, with 1 nm of spectral resolution. The methodological scheme
followed in this study is presented in Figure~\ref{fig-workflow}.

@@

\caption{\label{fig-VIP}Variable Importance of the Neural Network
Classifier for each vegetation class. The bigger the slice, the more
important is the variable to predict accuratly this class. The bottom
right plot shows the drone standardised reflectance spectra of each
class. Each slice represent the VIP of both RAW and standardised
reflectance combined.}

@@

The primary objective of this study was to develop a method for the
accurate classification of macrophytes on intertidal mudflats,
specifically focusing on distinguishing between Chlorophyceae (green
algae) and marine Magnoliopsida (seagrasses) using multispectral drone
data. The ability to differentiate between various types of vegetation,
as demonstrated in this study, plays a critical role in ecological
monitoring and management. By distinguishing between seagrasses and
macroalgae, our approach facilitates targeted conservation strategies,
enabling more effective preservation and restoration efforts in coastal
ecosystems. The discrimination of seagrasses from green macroalgae
presents significant challenges, primarily due to the shared pigment
composition between these taxa, as depicted in Figure~\ref{fig-Pigm} but
also by the frequent spatial mixing of these green macrophytes. This
similarity in pigment composition, including photosynthetic and
carotenoid pigments like chlorophyll b and c, fucoxanthin, and
neoxanthin, results in spectral signatures that are often
indistinguishable to non exper eyes. This spectral overlap is further
complicated by the intermingling of seagrasses and green macroalgae in
the same spatial locations, posing a significant challenge for remote
sensing aimed at accurately mapping and monitoring coastal ecosystems.
Our study addresses these complexities by utilizing high-resolution
drone imagery with 10 spectral bands. \citep{Davies2023} has shown that
having at least 8 spectral bands ranging between 500 nm to 850 nm with
including a band at 530 nm and another one at 730 nm is crucial to
accurately discriminate green macrophytes.

@@

Climate change, global warming, alien and invasive species development,
coastal erosion, sealevel rise, while further impact coastal ecosystems
in the future \citetext{\citealp{SCHIBALSKI2022101414}; \citealp[
]{holon2018predictive}; \citealp{marquet2024global}}. The demand for
meaningful and efficient monitoring methods for these habitats has
therefore never been greater \citetext{\citealp[
]{muller2018satellite}; \citealp[
]{villalobos2023remote}; \citealp{oiry2021using}}. Our findings,
particularly the improved discrimination of vegetation types, underscore
the potential of drone-based remote sensing to support diverse
applications, from biodiversity conservation to climate change
adaptation strategies\textbf{.} Employing traditional sampling methods
to monitor these ecosystems is time and resource-intensive, and the
findings are often difficult to scale up. Earth observation methods can
bridge this gap and meet the needs for monitoring coastal ecosystems.
\citep{papathanasopoulou2019satellite}. The retrieval of Essential
Biodiversity Variables (EBVs) and Essential Ocean Variables (EOVs)
through satellite observations is increasingly common, enabling
comprehensive monitoring of entire ecosystems over extended time periods
\citetext{\citealp[
]{ratnarajah2023monitoring}; \citealp{ZOFFOLI2020112020}}. The
significance of monitoring waters, particularly in light of the Water
Framework Directive (WFD ; \citep{WFD2000}), cannot be overstated in
environmental management and conservation efforts. The WFD mandates the
achievement and maintenance of ``good ecological status'' for all
European waters, which necessitates a comprehensive understanding and
monitoring of aquatic ecosystems, including coastal habitats like
seagrass \citetext{\citealp[ ]{foden2007angiosperms}; \citealp[
]{nordlund2024one}; \citealp{Zoffoli2021}}.
